---
title: "STA 207: Assignment II"
date: "Due 5 pm PST Feb 12th"
author: "Christina De Cesaris and student ID)"
output: html_document
---
***

**Instructions** You may adapt the code in the course materials or any sources (e.g., the Internet, classmates, friends). In fact, you can craft solutions for almost all questions from the course materials with minor modifications. However, you need to write up your own solutions and acknowledge all sources that you have cited in the Acknowledgement section. 

Failing to acknowledge any non-original efforts will be counted as plagiarism. This incidence will be reported to the Student Judicial Affairs. 

*** 

Two questions to be graded by the TA: 

***

A consulting firm is investigating the relationship between wages and some demographic factors. The file `Wage.csv` contains three columns, which are 

  - `wage`, the wage of the subject,
  - `ethnicity`, the ethnicity of the subject,
  - and `occupation`, the occupation of the subject. 


```{r,echo=T,results=F,message=F}
Wage=read.csv('Wage.csv');
library(gplots)
library(lme4)
attach(Wage)
```

***

(1) Write down a two-way ANOVA model for this data. For consistency, choose the letters from $\{Y,\alpha, \beta, \mu, \epsilon\}$ and use the factor-effect form. 

***

(2) Obtain the main effects plots and the interaction plot. Summarize your findings.
	
***
	
(3) Fit the ANOVA model described in Part 1. Obtain the ANOVA table and state your conclusions. Are the findings here consistent with your initial assessments from Part 2?

***

(4) Carry out a test to decide if the  effect of ethnicity is present on the full data set, at the significance level $\alpha=0.01$. 
	
***	

(5) For this part and the next, assume that the occupations have been selected randomly. Write down an appropriate ANOVA model that is additive in the factors and explain the terms in the model.
	
	

***

(6) Assuming that the model in Part 5 is appropriate, obtain an estimate of the proportion of variability that is due to variability in occupation.
	
*** 

## Acknowledgement {-}

## Session information {-}
```{r}
sessionInfo()
```