---
title: "STA 207: Assignment II"
date: "Due 5 pm PST Feb 12th"
author: "Christina De Cesaris; 913819951"
output: html_document
---


Two questions to be graded by the TA: 



A consulting firm is investigating the relationship between wages and some demographic factors. The file `Wage.csv` contains three columns, which are 

  - `wage`, the wage of the subject,
  - `ethnicity`, the ethnicity of the subject,
  - and `occupation`, the occupation of the subject. 


```{r,echo=T,results=F,message=F}
Wage=read.csv('Wage.csv');
library(gplots)
library(lme4)
attach(Wage)
```

***

(1) Write down a two-way ANOVA model for this data. For consistency, choose the letters from $\{Y,\alpha, \beta, \mu, \epsilon\}$ and use the factor-effect form. 

```{r Data Summary}

dim(Wage) # 534 instances (employees) by 3 features
which(is.na(Wage)) #no missing values
str(Wage)


table(Wage$ethn)
table(Wage$occ)


```
In this project, we define the two-way ANOVA model for predicting **wage** on **occupation** and **ethnicity** as:


$Y_{ijk} = \mu_{\cdot\cdot} + \alpha_i+\beta_j + (\alpha\beta)_{ij}+\epsilon_{ijk}, \ k=1,\ldots, 534, j=1,\ldots, 3, i=1,\ldots, 6,$ ,

where the index $i$ represents the occupation, and $j$ represent ethnicity. For example, if  $i=1$ and $j=2$, $Y_{ijk}$ would represent the wage of the $kth$ hispanic management employee. 

The term $\alpha\beta_{ij}$ for $i=1$ and $j=2$ would represent the interaction effect between hispanic ethnicities and management occupations. 


The error terms $\{\epsilon_{ijk}\}$ are i.i.d. $N(0,\sigma^2)$  



The total means among levels of both groups are defined as


$$
\mu_{\cdot \cdot} =\sum_{i=1}^a \sum_{j=1}^ b \mu_{ij}/(ab)$$
And the means for each group is:

$$\ \mu_{i\cdot} = \sum_{j=1}^3 \mu_{ij} /3 \\ \ \mu_{\cdot j}=\sum_{i=1}^6 \mu_{ij}/6 \\\mu_{\cdot \cdot} =\sum_{i=1}^6 \sum_{j=1}^ 3 \mu_{ij}/(ab)$$

In the case above, if $i=1$, $\mu_{1\cdot}$ would be the average wage for management positions. On the other hand if $j=2$, $\mu_{\cdot j}$ would represent the average wage for hispanic employees across all occupations. 


In practice, the factor effects form is preferred. We the factor effects for this model as: 

$$\alpha_i=\mu_{i\cdot} - \mu_{\cdot \cdot}\\ \beta_j=\mu_{\cdot j}-\mu_{\cdot\cdot}\\  (\alpha\beta)_{ij} =\mu_{ij}-\mu_{i\cdot}-\mu_{\cdot j}+\mu_{\cdot\cdot}$$


With the natural constraints:

$$\sum^6_{i=1}{\alpha_i} = \sum^3_{j=1} {\beta_j}=\sum_{i=1}^6 (\alpha\beta)_{ij}=\sum_{j=1}^3 (\alpha\beta)_{ij} =0\\$$


Our final factor effect model with constraints becomes:
$$Y_{ijk} = \mu_{\cdot\cdot} + \alpha_i+\beta_j + (\alpha\beta)_{ij}+\epsilon_{ijk}, \ k=1,\ldots, 534,  \ j=1,\ldots, 3, \ i=1,\ldots, 6 $$
For this Two-Way Anova model, I assume
  - Variance of $\epsilon_{ijk}$ error terms are equal
  - $\epsilon_{ijk}$ error terms are independent
  - $\epsilon_{ijk}$ error terms are normally distributed
  - There are no samples in our data that deviates vastly from the model (no outliers)
  - Our model is complete without any missing variables
  - There is equal variance across factor groups

***

(2) Obtain the main effects plots and the interaction plot. Summarize your findings.
	
	
```{r, fig.with=30}


# Main effect plot for ingredient 1
plotmeans(wage~occupation,data=Wage,xlab="Occupation",ylab="Wage", main="Main  effect, Occupation") 
# Main effect plot for ingredient 2
plotmeans(wage~ethnicity,data=Wage,xlab="Ethnicity",ylab="Wage", main="Main  effect, Ethnicity") 
#Interaction plot
interaction.plot(Wage$occup, Wage$ethnicity, Wage$wage, main="Interaction effect, Ethnicity and Occupation", xlab="Occupation",ylab="Wage", trace.label =  "Ethnicity", xtick = T)

```
TODO:**Summarize your findings**

***
	
(3) Fit the ANOVA model described in Part 1. Obtain the ANOVA table and state your conclusions. Are the findings here consistent with your initial assessments from Part 2?

```{r}
anova.1 = aov(wage~factor(occupation)*factor(ethnicity), data = Wage )
anova.1$coeff
summary(anova.1)
```

***

(4) Carry out a test to decide if the  effect of ethnicity is present on the full data set, at the significance level $\alpha=0.01$. 


**Test for the main effect of factor A.** Consider the following null and alternative hypotheses. 

$$
H_0: \beta_{i}=0 \ \ \forall i \ \ \  {\rm v.s.}\ \  \ H_a:\  {\rm not \ all \ } \beta_i \ {\rm are \ zero}.
$$




We have $F^*={\rm MSA}/{\rm MSE}$ that follows an F-distribution with $F(0.99, 2,516 )$. If $F^* > F(0.99, 2,516)$ then we rejuect the null hypothsis. 

$F^*=$
```{r}

round(summary(anova.1)[[1]][2,4],digits=3)

```

$F(0.01, 2,516 )$
```{r}
qf(0.99, 2,516 )

```
$2.159\ngtr 4.646516$ so $ F^* \ngtr F(0.99, 2,516)$ and therefore we do not reject the null hypothesis at significance level $\alpha=0.01$. 
	
***	

(5) For this part and the next, assume that the occupations have been selected randomly. Write down an appropriate ANOVA model that is additive in the factors and explain the terms in the model.
	Mixed effects model.
	
	
	
	When fixed effects and random effects are both present, our model takes the same form but there are a few more assumptions 
$$
Y_{ijk} =  \mu_{\cdot\cdot} + \alpha_i+\beta_j + (\alpha\beta)_{ij}+\epsilon_{ijk}, \ k=1,\ldots, 534, j=1,\ldots, 3, i=1,\ldots, 6,
$$
where (i) $\sum \alpha_i =0$, (ii) $\beta_j$ are i.i.d. $N(0,\sigma_{\beta}^2)$, (iii) $\sum_i  (\alpha\beta)_{ij} =0$ for any $j$, 
(iv) $ (\alpha\beta)_{ij} \sim N(0,(1-1/a)\sigma^2_{\alpha\beta}) $, (v) ${\rm cov}( (\alpha\beta)_{ij},  (\alpha\beta)_{i'j})= -\sigma^2_{\alpha\beta}/a$, (vi) ${\rm cov}( (\alpha\beta)_{ij},  (\alpha\beta)_{i'j'})=0$,  if $i\neq i'$ and $j\neq j'$, (vii)$\{\epsilon_{ijk}\}$ are i.i.d. $N(0,\sigma^2)$, and (viii)  $\{ \beta_j\}$, $\{(\alpha\beta)_{ij}\}$, $\{\epsilon_{ijk} \}$ are pairwise independent. 

In this model, we have 
1. $ \mathbb{E}[Y_{ijk}]=\mu_{\cdot\cdot}+\alpha_i$ 
2. ${\rm var}(Y_{ijk})=\sigma_{\beta}^2+(1-1/a)\sigma_{\alpha\beta}^2+\sigma^2$. 
3. ${\rm cov}(Y_{ijk},Y_{ijk'})=\sigma_{\beta}^2+(1-1/a)\sigma_{\alpha\beta}^2$ for $k\neq k'$. 
4. ${\rm cov}(Y_{ijk},Y_{ij'k'})=0$ for $j\neq j'$.
5. ${\rm cov}(Y_{ijk},Y_{i'jk'})=\sigma_{\beta}^2-1/a\sigma_{\alpha\beta}^2$ for $i\neq i'$.
6. ${\rm cov}(Y_{ijk},Y_{i'j'k'})=0$ when no indices are equal.


***

(6) Assuming that the model in Part 5 is appropriate, obtain an estimate of the proportion of variability that is due to variability in occupation.
	We can see that 
$\mathbb{E}[{\rm MSAB}]- \mathbb{E}[{\rm MSE}]=n\sigma^2_{\alpha\beta},  \mathbb{E}[{\rm MSB}]-\mathbb{E}[{\rm MSAB}]=n a \sigma_{\beta}^2$. We can define unbiased estimators of the variance components based on the mean squares. 

We have 
$\hat{H}= \sum c_i \bar{Y}_{i\cdot \cdot}$, ${\rm var}(\hat{H}) = \sum c_i^2(n\sigma_{\alpha\beta}^2+\sigma^2)/(nb)$, and $s^2(\hat{H})=\sum c_i^2 {\rm MSAB}/(nb)$. 
	

*** 

## Acknowledgement {-}

## Session information {-}
```{r}
sessionInfo()
```